### 2017-01-21 Thesis/latex/acknowledgements.tex
New git repo

\chapter{Acknowledgements}

First of all I would like to thank Henrik Kalisch and Magnus Sv\"ard who carried out the selection process of candidates for the PhD position. I am sincerely thankful to Professor Henrik Kalisch, my main supervisor, for sharing knowledge and guiding me, for his enthusiasm, support and encouragement. The thesis would not have been possible without you.

In the course of the doctorate studies, I have had fruitful collaboration with Olivier Verdier, Evgueni Dinvay, Denys Dutykh, Ben Segal and others. I would like to thank them for productive discussions, advices on research questions and some inspiring ideas we came up with.

As a PhD Candidate I have been lucky to meet my colleagues and fellow students at the Department of Mathematics. They have always provided a friendly and supportive environment. I appreciate everything we have experienced during the time of our studies.

I would like to express my gratitude to my family for their continuous support and encouragement over the years of my studies. I am also thankful for all of my teachers and professors from schools and universities I studied at.

I am heartily grateful for my wife Assem. You have always been there for me. Without your love, support and understanding I would not have been able to go that far.