### 2017-01-21 Thesis/latex/outline.tex

\chapter{Outline}

This thesis is organised in the following way. Part I contains general theoretical background on nonlinear wave models as well as description of methods used to solve the equations involved. Some properties of solutions to the equations and a summary of results are also given in the first part. Part 2 consists of the research papers that present scientific results in detail.

\section*{List of reseach papers included in Part II}

\textbf{Paper A:}

\noindent Moldabayev, D., Kalisch, H., Dutykh, D.: \textit{The Whitham equation as a model for surface water waves}, Phys. D~ \textbf{309}, 99–107 (2015),

\noindent http://dx.doi.org/10.1016/j.physd.2015.07.010.

\vspace{1em}

\noindent \textbf{Paper B:}

\noindent Dinvay, E., Moldabayev, D., Dutykh, D., Kalisch, H.:  \textit{The Whitham equation with surface tension}, Nonlinear Dynamics, 1--14 (2017),

\noindent http://dx.doi.org/10.1007/s11071-016-3299-7.

\vspace{1em}

\noindent \textbf{Paper C:}

\noindent Henrik Kalisch, Daulet Moldabayev, Olivier Verdier: \textit{A numerical study of nonlinear dispersive wave models with SpecTraVVave}, \textcolor{red}{specify status of the paper.}

\vspace{1em}

\noindent \textbf{Paper D:}

\noindent Benjamin Segal, Daulet Moldabayev, Henrik Kalisch, Bernard Deconinck: \textit{Explicit solutions for a long-wave model with constant vorticity}, submitted to European Journal of Mechanics - B/Fluids.