### 2017-01-21 Thesis/latex/papers.tex

%\part{Scientific results}

%--------------------------------------------------------------------------------------------------
% Paper A
%--------------------------------------------------------------------------------------------------
\chapter*{Paper A}

\addcontentsline{toc}{chapter}{Paper A The Whitham equation as a model for surface water waves}
\section*{The Whitham equation as a model for surface water waves}

\noindent Daulet Moldabayev, Henrik Kalisch, Denys Dutyh \\

\noindent \textit{Physica} D \textbf{309}, (2015)

\noindent http://dx.doi.org/10.1016/j.physd.2015.07.010
\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/paper-1/page_}{9}

%--------------------------------------------------------------------------------------------------
% Paper B
%--------------------------------------------------------------------------------------------------

\chapter*{Paper B}

\addcontentsline{toc}{chapter}{Paper B The Whitham equation with surface tension}
\section*{The Whitham equation with surface tension}

\noindent Evgueni Dinvay, Daulet Moldabayev, Denys Dutykh, Henrik Kalisch\\

\noindent \textit{Nonlinear Dynamics}, 1--14 (2017)

\noindent http://dx.doi.org/10.1007/s11071-016-3299-7

\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/paper-2/page_}{14}

%--------------------------------------------------------------------------------------------------
% Paper C
%--------------------------------------------------------------------------------------------------

\chapter*{Paper C}

\addcontentsline{toc}{chapter}{Paper C A numerical study of nonlinear dispersive wave models with SpecTraVVave}
\section*{A numerical study of nonlinear dispersive wave models with SpecTraVVave}

\noindent Henrik Kalisch, Daulet Moldabayev, Olivier Verdier\\

%\noindent \textit{Physical Review} B, \textbf{76}, 035303 (2007)
\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/paper-3/page_}{27}

%--------------------------------------------------------------------------------------------------
% Paper D
%--------------------------------------------------------------------------------------------------

\chapter*{Paper D}

\addcontentsline{toc}{chapter}{Paper D Explicit solutions for a long-wave model with constant vorticity}
\section*{Explicit solutions for a long-wave model with constant vorticity}

\noindent Benjamin Segal, Daulet Moldabayev, Henrik Kalisch, Bernard Deconinck\\

%\noindent \textit{Physical Review} B, \textbf{76}, 035303 (2007)
\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/paper-4/page_}{19}