### 2017-01-21 Thesis/latex/scientific_environment.tex

\chapter{Scientific environment}

This is where you write the name of the subject/professional milieu involved in the study, faculty(faculties)/institute(s)/centre(s)/research groups/school of research.

Contractual joint venture partners should also be mentioned here. If these are represented by graphic symbols (logo etc.) the it should be on this page.

NB! Most organisations have guidelines concerning the use of their logo. These guidelines regulate who can use the logo, when the logo can be used and how. This is quite relevant when a logo is used together with information connected to other institutions/organisations. (A typical example would be rules concerning space between other written or graphic content.)

Remember to find out which guidelines apply and ensure that they are followed!
Most organisations will be able to provide a digital version of their logo in a suitable format.