### 2017-01-21 Thesis/papers.tex

%\chapter*{Scientific results}
%\addcontentsline{toc}{chapter}{Scientific results}

%--------------------------------------------------------------------------------------------------
% Paper I
%--------------------------------------------------------------------------------------------------
\chapter*{Paper A}
\addcontentsline{toc}{chapter}{Paper A - Linear and nonlinear convection in porous media between coaxial cylinders}
\section*{Linear and nonlinear convection in porous media between coaxial cylinders}
%\addcontentsline{toc}{section}{Linear and nonlinear convection in porous media between coaxial cylinders}

\noindent C. Bringedal, I. Berre, J.M. Nordbotten, D.A.S. Rees\\

\noindent \textit{Physics of Fluids} \textbf{23}, 9, (2011). \\ http://dx.doi.org/10.1063/1.3637642
\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
%\newgeometry{hmargin=0.5cm,vmargin={0.5cm}}
\includepaperpages{../papers/Paper1/page_}{11}
%\includepdf[pages={2-12},lastpage=12]{../papers/Paper1}
%\restoregeometry

%\chapter*{Paper B}
%\section{Influence of natural convection in a porous medium when producing from borehole heat exchangers}
\chapter*{Paper B}
\addcontentsline{toc}{chapter}{Paper B - Influence of natural convection in a porous medium when producing from borehole heat exchangers}  
\section*{Influence of natural convection in a porous medium when producing from borehole heat exchangers}
%\addcontentsline{toc}{section}{Influence of natural convection in a porous medium when producing from borehole heat exchangers}

\noindent C. Bringedal, I. Berre, J.M. Nordbotten, \\

\noindent \textit{Water Resources Research} \textbf{49}, 8, (2013). \\ http://dx.doi.org/10.1002/wrcr.20388

\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/Paper2/page_}{12}

\chapter*{Paper C}
\addcontentsline{toc}{chapter}{Paper C - An Approach for Investigation of Geochemical Rock-Fluid Interactions}  
\section*{An Approach for Investigation of Geochemical Rock-Fluid Interactions}
%\addcontentsline{toc}{section}{An Approach for Investigation of Geochemical Rock-Fluid Interactions}

\noindent C. Bringedal, I. Berre, F.A. Radu, \\

\noindent \textit{Proceedings, Thirty-Ninth Workshop on Geothermal Reservoir Engineering}, Stanford University, (2014).

\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/Paper3/page_}{10}

\chapter*{Paper D}
\addcontentsline{toc}{chapter}{Paper D - A model for non-isothermal flow and mineral precipitation and dissolution in a thin strip}  
\section*{A model for non-isothermal flow and mineral precipitation and dissolution in a thin strip}
%\addcontentsline{toc}{section}{A model for non-isothermal flow and mineral precipitation and dissolution in a thin strip}

\noindent C. Bringedal, I. Berre, I.S. Pop, F.A. Radu, \\

%\noindent \textit{Journal of Computational and Applied Mathematics}, \\http://dx.doi.org/10.1016/j.cam.2014.12.009, (2014).

\noindent \textit{Journal of Computational and Applied Mathematics} \textbf{289}, (2015)\\ http://dx.doi.org/10.1016/j.cam.2014.12.009

\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/Paper4/page_}{10}

\chapter*{Paper E}
\addcontentsline{toc}{chapter}{Paper E - Upscaling of non-isothermal reactive porous media flow with changing porosity}  
\section*{Upscaling of non-isothermal reactive porous media flow with changing porosity}
%\addcontentsline{toc}{section}{Upscaling of non-isothermal reactive porous media flow with changing porosity}

\noindent C. Bringedal, I. Berre, I.S. Pop, F.A. Radu, \\

\noindent \textit{Transport in Porous Media} (2015)\\ http://dx.doi.org/10.1007/s11242-015-0530-9
%\noindent \textit{Submitted to Transport in Porous Media}

\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/Paper5/page_}{26}

\chapter*{Paper F}
\addcontentsline{toc}{chapter}{Paper F - Upscaling of non-isothermal reactive porous media flow under dominant P\'{e}clet number: the effect of changing porosity}  
\section*{Upscaling of non-isothermal reactive porous media flow under dominant P\'{e}clet number: the effect of changing porosity}
%\addcontentsline{toc}{section}{Upscaling of non-isothermal reactive porous media flow under dominant P\'{e}clet number: the effect of changing porosity}

\noindent C. Bringedal, I. Berre, I.S. Pop, F.A. Radu, \\

\noindent \textit{Submitted}% to SIAM Multiscale Modeling $\&$ Simulations}.

\cleardoublepage

% This will include the bursted pdf. Remember to update the number of pages
% (last parameter) to the actual number of pages
\includepaperpages{../papers/Paper6/page_}{30}