### 2017-02-19 Babenko/BabenkoEquation/lecture/statement.tex

\section{The Babenko equation}
\subsection{Statement of the problem}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{The Babenko equation: statement of the problem}
Equations of the following form are considered:

\begin{equation}
u_t+ f(u)_x + \Lop u_x=0 
\end{equation}

where $f(\eta)$ is a non-linear flux function and $\mathcal{L}$ is a linear dispersion operator:

$$\widehat{\Lop u}(k)=\alpha(k)~\widehat{u}(k). $$

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%---------------------------------------------------------------------%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{The Babenko equation: statement of the problem}
Equations of the following form are considered:

   \onslide<1-2>{The Whitham equation:

\begin{equation}
u_t+\frac{3}{2}u u_x+K\ast u_x=0. \label{Whitham}
\end{equation}

$$
\widehat{K\ast u}(k) = \sqrt{\frac{\tanh(k)}{k}}~\widehat{u}(k).
$$}
   \onslide<2-2>{The KdV equation:
\begin{equation}
u_t+\frac{3}{2}u u_x+ u_x + \frac{1}{6} u_{xxx}=0.
\end{equation}}
\end{frame}