### 2017-03-26 Thesis/latex/preface.tex

\chapter{Preface}

This dissertation is submitted as a partial fulfillment of the requirements for the degree Doctor of Philosophy (PhD) at the Department of Mathematics, University of Bergen. The research was
supported by the Research Council of Norway on grant no. 213474/F20.

The advisory committee has consisted of Henrik Kalisch and Guttorm Alendal (University of Bergen, Bergen, Norway),  Dmitry Pelinovsky (McMaster University, Hamilton, Ontario, Canada) and Angel Dur\'an (Universidad de Valladolid, Valladolid, Spain).